### 2021-03-24 input/proposal.tex
about implem

Each process is stateless, works autonomously and asynchronously, in a
peer-to-peer fashion. Each process broadcasts each change to its
neighborhood, for the state of its neighbors may depend on this change
as well. Yet, the traffic generated to to adapt the scope of
S-broadcast remains bounded, for it uses the principles of scoped
broadcast itself. Messages that carry changes travel through the
network depending on partition of processes, stopping as soon as they
encounter uninterested processes. We demonstrate that \NAME guarantees
consistent partitioning despite different order in message deliveries
from one process to another.

@@

\begin{theorem}[Dynamic consistent partitioning]
  Eventual consistent partitioning in presence of dynamic sources
  requires
  \begin{inparaenum}[(i)]
  \item eventual purging of stale notifications, and
  \item eventual delivery of bests partitions.
  \end{inparaenum}
\end{theorem}

@@

Algorithm~\ref{algo:adddelundo} provides the instructions that
implement dynamic consistent partitioning:
\begin{inparaenum}[(i)]
\item most of the time, it propagates $\alpha$'s and $\delta$'s when
  the partitions allow it;
\item sometimes, it detects possible inconsistent partitioning and
\item solves it using propagation trees (as opposed to propagation
  graphs);
\item when required, it triggers competitions among neighboring
  partitions.
\end{inparaenum}

@@

\noindent To implement last win order, each process maintains a vector
of versions that associates to each known source, or has-been source,
its known local counter. It enables processes to deliver only
up-to-date $\alpha$'s that may improve their best known partition (see
Line~\ref{line:detectA}), but also to detect possible inconsistent
partitioning (see Line~\ref{line:detectB}) as labeled by III in
Figure~\ref{fig:proofB}. From a global perspective, it ensures
monotonic increase of knowledge despite the non-monotonic number of
sources. In turns, it contributes to convergence and termination, for
corresponding $\alpha$ and $\delta$ cannot follow each other in an
infinite loop.

\noindent Upon possible inconsistency, the detecting process must
remove all $\alpha$ messages originating from it, as labeled by IV in
Figure~\ref{fig:proofB}. Contrarily to delete operations, this applies
to a sub-partition of the current partition, and cannot rely on local
counters to operate. Therefore, we implement this behavior by
piggybacking paths in messages.  \TODO{Paths and \NAME synergies well,
  for paths tend to be smaller as the number of sources increase in
  the system.}

\noindent Deletes must trigger competition amongst neighboring
partitions. These \TODO{add} messages operate normally and fill gaps
left open by deletes.

@@

\paragraph{Dynamic network.}
Adding new communication links to the network may create shortcuts
between processes. Both processes must send their current best
partition to each other. Upon receipt, they act normally: if a process
finds out that the received partition is closer than its current one,
it delivers it which in turns also triggers another competition
amongst neighbors due to forwarding.

\noindent Joining the network is equivalent to add as many
communication links as necessary between the joining process and its
new neighbors.

When removing a communication link between two processes does not
break any active path, because neither distances of processes depend
on the other, then nothing needs to be done. \NAME has no overhead.
Unfortunately, when a process' distance depends on the other process,
the protocol becomes much more complex. Indeed, this requires to
\TODO{undo} all add messages originated from this process. A message
must convey the fact that
\begin{inparaenum}[(i)]
\item an edge at a particular process has been removed, and
\item the distance that has been delivered by a process comes from
  this particular process.
\end{inparaenum}

\noindent \NAME handles this as a particular case of partition within
current partition: it contains to affected regions the traffic
generated to patch affected regions. In order to know if it is
affected by the \TODO{undo} operation, using distances already
piggybacked in messages is not sufficient.  Each process must know the
path taken by delivered message. Since we do not beforehand which
communication links will crash and which messages will be affected by
undos, each message has to carry its path along forwarding. This also
allows processes to remove loops, for they cannot rely on version
number (that remains unchanged) for this operation.

\noindent Instead of version number, one could use such paths to remove loops
due to \TODO{add and del}, however this may be much more costly in
terms of generated traffic. Version numbers guarantee that only one
delete is delivered and all prior inserts are then discarded, while
using paths, processes may deliver multiple inserts and deletes before
reaching termination. \TODO{maybe clearer with figures.} As
consequence, \NAME still makes use of vector of versions.

\noindent Although costly, piggybacking paths with logical partitioning synergies
well, for the size of these paths decreases quickly as the number of
partitions grows.

\paragraph{Bloom filters.}
We also propose an implementation that further decreases generated
traffic and improves on anonymity~\cite{whitaker2002forwarding}.
\NAME makes extensive use of paths to enable deleting messages without
incrementing local counters. Messages carry their respective path and
processes detect when such messages have been looping when they carry
their identity (see Lines~\ref{line:notloopingA} and
\ref{line:notloopingB} of Algorithm~\ref{algo:adddelundo}). Since
membership is all that matters, we propose to trade vectors of
identities for Bloom filters~\cite{almeida2007scalable}. Processes
know with high probability if they already received and forwarded each
message without knowing the identity of all processes that received
it.  False positive probability only incurs premature stopping of
broadcast messages and does not invalidate the delete of specific
messages.

>>>>

Each process works autonomously and asynchronously, in a peer-to-peer
fashion. Each process broadcasts each change to its neighborhood, for
the state of its neighbors may depend on this change as well. Yet, the
traffic generated to to adapt the scope of scoped broadcast remains
bounded, for it uses the principles of scoped broadcast
itself. Messages that carry changes travel through the network
depending on partition of processes, stopping as soon as they
encounter uninterested processes. We demonstrate that \NAME guarantees
consistent partitioning despite different order in message deliveries
from one process to another.

@@

\begin{theorem}[\label{theo:dcp}Dynamic consistent partitioning]
  Eventual consistent partitioning in presence of dynamic sources
  requires
  \begin{inparaenum}[(i)]
  \item eventual purging of stale notifications, and
  \item eventual delivery of bests partitions.
  \end{inparaenum}
\end{theorem}

@@

Algorithm~\ref{algo:adddelundo} provides the instructions that
implement dynamic consistent partitioning:
\begin{inparaenum}[(i)]
\item \label{algo:most} most of the time, it propagates $\alpha$'s and
  $\delta$'s when the partitions allow it;
\item \label{algo:sometimes} sometimes, it detects possible
  inconsistent partitioning and
\item \label{algo:solves}solves it using propagation trees (as opposed
  to propagation graphs);
\item \label{algo:competition} when required, it triggers competitions
  among neighboring partitions.
\end{inparaenum}
While (\ref{algo:most})-(\ref{algo:solves}) tackle the eventual purging of
stale notifications, (\ref{algo:competition}) tackle the eventual
delivery of the best up-to-date partitions.

@@

To implement last win order, each process maintains a vector of
versions that associates to each known source, or has-been source, its
known local counter. It enables processes to deliver only up-to-date
$\alpha$'s that may improve their best known partition (see
Line~\ref{line:detectA}), but also to detect possible inconsistent
partitioning (see Line~\ref{line:detectB}) as labeled by III in
Figure~\ref{fig:proofB}. From a global perspective, it ensures
monotonic increase of knowledge despite the non-monotonic number of
sources. In turns, it contributes to convergence and termination, for
corresponding $\alpha$ and $\delta$ cannot follow each other in an
infinite loop.

Implementing the propagation of deletes is no different from the
propagation of adds as hinted in Figure~\ref{fig:proof} with Labels I
and I'. Messages and comparisons include the local counter of the
source. When a process receives a better up-to-date partition, it
delivers and forwards it. When a process receives a delete
notification about its partition, it delivers and forwards it. It is
worth noting that such $\delta$'s apply to the whole partition. They
travel in the propagation graph of their partition using all
communication links, allowing processes to change partition as soon as
possible.

Vectors of versions also allow each process to detect possible last
win order violations. Upon possible inconsistency, the detecting
process must remove all $\alpha$ messages originating from it, as
labeled by IV in Figure~\ref{fig:proofB}. Contrarily to delete
operations, this applies to a sub-partition of the current partition,
and cannot rely on local counters to
operate. Algorithm~\ref{algo:adddelundo} implements this behavior by
piggybacking paths in messages. Paths guarantee that such $\delta$
messages apply to all and only $\alpha$'s that followed an identical
path. $\delta$'s propagate in the propagation tree rooted at the
detecting process. Local states of processes remain unchanged, but
paths in messages monotonically grow. In turns, $\alpha$ and $\delta$
cannot follow each other in an infinite loop, for a process knows when
it already received a forwarded a message (see
Lines~\ref{line:notloopingA} and \ref{line:notloopingB}).

\noindent It is worth noting that scoped broadcast and the
piggybacking of paths synergies well: Paths tend to be smaller as the
number of sources in the system increases.

Finally, as stated in Theorem~\ref{theo:dcp}, dynamic consistent
partitioning not only requires the removal of all stale partitions,
but also the retrieval of the best up-to-date partitions. In that
regard, $\delta$'s have dual use: since they already reach borders of
partitions when they remove stale control informations, they also
trigger a competition when reaching such bordering processes (see
Line~\ref{line:compete}). This simply consists in sending the current
partition through the communication link from which the process
received the $\delta$. Upon receipt of this answer, processes act
normally by propagating their changes when they improve. These adds
fill the gaps left open by deletes.

@@

\paragraph{Dynamic network.}
Algorithm~\ref{algo:edges} provides the instructions of \NAME that
enable dynamic consistent partitioning in dynamic networks where
processes can join or leave the system without
notification. \TODO{More.}

%% Adding new communication links to the network may create shortcuts
%% between processes. Both processes must send their current best
%% partition to each other. Upon receipt, they act normally: if a process
%% finds out that the received partition is closer than its current one,
%% it delivers it which in turns also triggers another competition
%% amongst neighbors due to forwarding.

%% \noindent Joining the network is equivalent to add as many
%% communication links as necessary between the joining process and its
%% new neighbors.

%% When removing a communication link between two processes does not
%% break any active path, because neither distances of processes depend
%% on the other, then nothing needs to be done. \NAME has no overhead.
%% Unfortunately, when a process' distance depends on the other process,
%% the protocol becomes much more complex. Indeed, this requires to
%% \TODO{undo} all add messages originated from this process. A message
%% must convey the fact that
%% \begin{inparaenum}[(i)]
%% \item an edge at a particular process has been removed, and
%% \item the distance that has been delivered by a process comes from
%%   this particular process.
%% \end{inparaenum}

\paragraph{Another trade-off.}
We propose an implementation that further decreases generated traffic
and improves on anonymity~\cite{whitaker2002forwarding}.  \NAME makes
extensive use of paths to enable deleting messages without
incrementing local counters. Messages carry their respective path and
processes detect when such messages have been looping when they carry
their identity (see Lines~\ref{line:notloopingA} and
\ref{line:notloopingB} of Algorithm~\ref{algo:adddelundo}). Since
membership is all that matters, we propose to trade vectors of
identities for Bloom filters~\cite{almeida2007scalable}. Processes
know with high probability if they already received and forwarded each
message without knowing the identity of all processes that received
it.  False positive probability only incurs premature stopping of
broadcast messages and does not invalidate the delete of specific
messages.
